FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-08-02 paper.tex
Fix typos; add paragraph indent so it doesn't look awful
Write a bunch about injection decision

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}
\usepackage{amsmath}
\usepackage{cleveref}

@@

% If using Emacs, put cursor on last closing paren and call "eval-last-sexp" (usually `C-x C-e') to regenerate
%% (save-excursion
%%  (let* ((toc (funcall imenu-create-index-function))
%%         (headers (mapcar (pcase-lambda (`(,str . ,loc)) str)
%%                          toc)))
%%    (beginning-of-buffer)
%%    (search-forward (concat "%% " "--- autogenerated TOC ---"))
%%    (next-line)
%%    (beginning-of-line)
%%    (let ((start (point)))
%%      (search-forward (concat "%% " "--- end autogenerated TOC ---"))
%%      (beginning-of-line)
%%      (delete-region start (point))
%%      (goto-char start))
%%    (cl-loop for header in headers
%%             do (insert (format "%% %s\n" header)))
%%    (next-line)))

%% --- autogenerated TOC ---
% --
% --
% <<header.tex
% Introduction
% Floating Point Exception Primer
% . Where do \NaN{}s come from?
% . . TODO good quote, try to work this in
% . Infinity
% . To Kill a Floating Point
% \FlowFPX{}
% . \FT{}
% . . Catching a \NaN{} Killer
% . . \NaN{} propagation monitoring
% . . \NaN{} injection
% . Visualization
% . \GPUFPX{}
% Case Studies
% . Surprise \NaN{}s: ShallowWaters
% . . CSTGs help us get a bigger picture
% . \NaN{} fuzzing to find a kill: OrdinaryDiffEq
% . Finch
% . A Bug-Hunting Safari
% . \GPUFPX{}: Oceananigans
% Related Work
% Discussion
% . Tracking more than \NaN{}s and \Inf{}s
% . Enhanced Fuzzing
% . Static FPX, Gradual Types
% Acknowledgments
% <<bib.tex.tex
% --
%% --- end autogenerated TOC ---

@@

@@

\begin{abstract}
  Reliable numerical computations are central to high-performance computing and machine learning.
  We present FlowFPX: a Julia-based tool for tracking the onset and flow of floating-point exceptions.
  FlowFPX's design exploits Julia's operator overloading to trace exception flows and even inject exceptions to accelerate testing.
  We present intuitive visualizations of summarized exception flows including how they are generated, propagated and killed, thus helping with debugging and repair.
\end{abstract}

@@

\subsubsection{Tracking Excptional Values}

@@

\begin{lstlisting}[language = Julia]
function Base.+(x::TrackedFloat64, y::TrackedFloat64)
  result = x.val + y.val
  check_error(+, result, x.val, y.val)
  TrackedFloat64(result)
end
\end{lstlisting}

@@

\begin{tabular}{ll}
  Arguments & Result & Event Type \\
  \hrule
  No \NaN{} & \NaN{} & Gen \\
  \NaN{} & \NaN{} & Prop \\
  \NaN{} & No \NaN{} & Kill \\
\end{tabular}

@@

The exception tracking infrastructure that we have gives us a simple way to inject exceptional values wherever we want in our code.
We have a global configuration system that allows us to turn on fuzzing.
Then, each time we intercept a \fp{} operation, we check to see if we should inject a \NaN{}.\footnote{At time of writing, \FT{} only injects \NaN{}, but it is trivial to extend the set of values to inject.}
Our full decision process is as follows:

\begin{enumerate}
\end{enumerate}

@@

% Local Variables:
% jinx-local-words: "CFL CSTG's CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq RxInfer ShallowWaters TrackedFloat isnan"
% End:

>>>>

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{microtype}
% \usepackage{fontspec}

@@

% If using Emacs, put cursor on last closing paren and call "eval-last-sexp" (usually `C-x C-e') to regenerate
% (save-excursion
%  (let* ((toc (funcall imenu-create-index-function))
%         (headers (mapcar (pcase-lambda (`(,str . ,loc)) str)
%                          toc)))
%    (beginning-of-buffer)
%    (search-forward (concat "%% " "--- autogenerated TOC ---"))
%    (next-line)
%    (beginning-of-line)
%    (let ((start (point)))
%      (search-forward (concat "%% " "--- end autogenerated TOC ---"))
%      (beginning-of-line)
%      (delete-region start (point))
%      (goto-char start))
%    (cl-loop for header in headers
%             do (insert (format "%% %s\n" header)))
%    (next-line)))

%% --- autogenerated TOC ---
% juliacon
% document
%   Introduction
%   Floating Point Exception Primer
%     Where do \NaN{
%       TODO good quote, try to work this in
%     Infinity
%     To Kill a Floating Point
%   \FlowFPX{
%     \FT{
%       Tracking Excptional Values
%       Fuzzing
%     Visualization
%     \GPUFPX{
%   Case Studies
%     Surprise \NaN{
%       CSTGs help us get a bigger picture
%     \NaN{
%     Finch
%     A Bug-Hunting Safari
%     \GPUFPX{
%   Related Work
%   Discussion
%     Tracking more than \NaN{
%     Enhanced Fuzzing
%     Static FPX, Gradual Types
%   Acknowledgments
% document
%% --- end autogenerated TOC ---

@@

% Misc typographical tweaks.
\setlength{\parindent}{1.2em}
% \setmonofont{JuliaMono}[Extension = .ttf, Path = ./, UprightFont = *-Regular, ItalicFont = *-Italic]

@@

\begin{abstract}
  Reliable numerical computations are central to high-performance computing and machine learning.
  We present FlowFPX: a tool set for debugging floating-point code.
  FlowFPX is composed of two primary parts: FloatTracker, a Julia-based tool for tracking the onset and flow of floating-point exceptions; and CSTG, a stack trace visualizer.
  FlowTracker's design exploits Julia's operator overloading to trace exception flows and even inject exceptions to accelerate testing.
  We present intuitive visualizations of summarized exception flows including how they are generated, propagated and killed, thus helping with debugging and repair.
\end{abstract}

@@

\subsubsection{Tracking Exceptional Values}

@@

\begin{lstlisting}[language = Julia]
function +(x::TrackedFloat64, y::TrackedFloat64)
  result = x.val + y.val
  check_error(+, result, x.val, y.val)
  TrackedFloat64(result)
end
\end{lstlisting}

@@

\begin{center}
  \begin{tabular}{ccc}
    Arguments & Result & Event Type \\
    \hline
    No \NaN{} & \NaN{} & Gen \\
    \NaN{} & \NaN{} & Prop \\
    \NaN{} & No \NaN{} & Kill \\
  \end{tabular}
\end{center}

@@

The exception tracking infrastructure that we have gives us a simple way to inject exceptional values wherever we want in our code.
We have a global configuration system that allows us to turn on fuzzing and configure where, when, and how many \NaN{}s\footnote{At time of writing, \FT{} only injects \NaN{}, but it is trivial to extend the set of values to inject.} to inject.

\FT{} exposes these configuration options:

\begin{description}
\item[\texttt{active::Bool}] Inject only if this is true.
\item[\texttt{n\_inject::Int64}] Inject at most this many \NaN{}s.
\item[\texttt{odds::Int64}] Inject if \texttt{rand(odds) == 1}.
\item[\texttt{functions::Array}] Inject if we are within the dynamic extent of any of these functions.
\item[\texttt{libraries::Array}] Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
\item[\texttt{replay::String}] Name of a file to read an injection recording from.
\item[\texttt{record::String}] Name of a file to write injection events to for later replay.
\end{description}

Then, each time we intercept a \fp{} operation, we check to see if we should inject a \NaN{}.
\cref{fig:injection_decision} illustrates how we decide whether or not to inject a \NaN{}, which is as follows:

\begin{enumerate}
  \item Are we replaying an injection recording? If yes, inject iff the recording says to inject.
  \item Is injection turned on? Don't inject if no.
  \item Are there any NaNs left for us to inject? Don't inject if we're out.
  \item Roll a \texttt{odds}-sided die---if we get anything other than 1, don't inject.
  \item Check the stack trace: are we within the dynamic extent of any of the functions mentioned in \texttt{functions}? I.e. if we have a set of functions that we have expressed interest in via the \texttt{functions} configuration option, do any of them show up anywhere on the stack? If we have some functions, but none of them are on the stack, don't inject.
  \item Check the top-most stack frame: (modulo the stack frames incurred by \FT{} intercepting the function call) is it in the list of modules mentioned in \texttt{modules}? If we have some modules we're interested in, but the top frame is \emph{not} from the set of interesting modules, don't inject.
  \item If we've made it this far, inject and decrement \texttt{n\_inject}.
\end{enumerate}

In our experience, requesting the stack trace from Julia is the most expensive part of this operation.
Therefore, we defer all the way until step 5 to request a stack trace.
This expensive region is show with the purple box in \cref{fig:injection_decision}.

\begin{figure}
  \centering
  \includegraphics[width=3in]{./fig/injection_decision_white_labels.png}
  \caption{Decision flowchart for injecting NaNs. Purple box indicates expensive checks.}
  \label{fig:injection_decision}
\end{figure}

@@

% Local Variables:
% jinx-local-words: "CFL CSTG's CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
% End:
